On using the decision trees to identify the local extrema in parallel global optimization algorithms

### 2023-02-01 2023/PAVT/ParallelDecisionTree.tex
minor corrections before submitting

\section{The algorithms used}\label{SecA}

@@

The algorithm stops in the case if the condition \(\Delta_{t_j} < \varepsilon\) is satisfied for even a  single value $t_j,\ 1\le\ j\le\ p$ from (\ref{agp4_R_sort}). This stop criterion (along with the criterion limiting the number of executed iterations usual for the  iteration methods) is used in the applied optimization problems, in which the global minimum point $x^*$ is unknown a priori.

@@

After completing a certain number of trials (for example, $100\ \ast\ N$), let us use all accumulated  points to initialize a suitable data structure and training the decision tree on its base. To make the  determining the adjacent points easier, we construct a uniform grid with certain step in each coordinate.  Next, we compute the values of the approximation of the objective function in these points  using the decision tree. Now, having a uniform grid of points and knowing the function values in each  point, let us find the closest point to the initial one (from the viewpoint of the Euclidian distance). This  closest point is a projection of the initial point onto the uniform grid, on which its neighbors can be  determined easily. Since the decision tree constructs the approximation of the initial function (even  piecewise constant), we can estimate the function values in the areas, in which the trials haven't been  executed earlier.

@@

	\item If the decision tree is used for the first time, construct a uniform grid

	\begin{displaymath} 
		Y'=\{ y'\in\ R^N:\ a_i\le\  {y'_i}^k \le\ b_i,\ 1\le\ i\le\ N,\ 1\le k\le\ \sqrt[N]{300}  \} 
	\end{displaymath}

>>>>

\section{Algorithms}\label{SecA}

@@

The algorithm stops in the case if the condition \(\Delta_{t_j} < \varepsilon\) is satisfied for even a  single value $t_j,\ 1\le\ j\le\ p$, from (\ref{agp4_R_sort}). This stop criterion (along with the criterion limiting the number of executed iterations usual for the  iteration methods) is used in the applied optimization problems, in which the global minimum point $y^*$ is unknown a priori.

@@

After completing a certain number of trials (for example, $100\cdot N$), let us use all accumulated  points to initialize a suitable data structure and training the decision tree on its base. To make the  determining the adjacent points easier, we construct a uniform grid with certain step in each coordinate.  Next, we compute the values of the approximation of the objective function in these points  using the decision tree. Now, having a uniform grid of points and knowing the function values in each  point, let us find the closest point to the initial one (from the viewpoint of the Euclidian distance). This  closest point is a projection of the initial point onto the uniform grid, on which its neighbors can be  determined easily. Since the decision tree constructs the approximation of the initial function (even  piecewise constant), we can estimate the function values in the areas, in which the trials haven't been  executed earlier.

@@

	\item If the decision tree is used for the first time, construct a uniform grid
	\begin{displaymath} 
		Y'=\{ y'\in R^N:\ a_i\le y'_i \le b_i,\ 1\le i\le N \},
	\end{displaymath} 
	where number of the grid nodes is a parameter of the method.
